Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-28 sports_analytics.Rmd
first draft of #13

## Competitions

- Kaggle March Madness [@lopez2015building] (jqas special issue: https://www.degruyter.com/journal/key/jqas/11/1/html)
- Big Data Bowl (special issue https://www.degruyter.com/journal/key/jqas/16/2/html)
- Big Data Cup
- SABR Case competition

## Conferences

- NESSIS/CASSIS
- Sloan
- CMU
- UConn sports analytics

@marchi2018analyzing

@R-Lahman, @R-pitchRx, @R-retrosheet

@@

- How significant is the element of chance in a particular sport? 
- Given that we know how the best teams are, who are the best players and how can we quantify their relative contributions? 
- What combinations of players work best together in a particular sport?

>>>>

Public research in sports analytics is driven in part by several notable competitions. 
These competitions have been an important source of new ideas and have contributed to the diversification of the field by breaking down barriers to entry.

In 2014, the *Journal of Quantitative Analysis in Sports* teamed up with Kaggle to create a March Madness competition to predict the winners of the NCAA men's basketball tournament. 
243 entrants competed for the $15,000 cash prize by submitting predicted probabilities for every possible pairwise matchup among the 68 teams in the tournament [@GlickmanSonas2015]. 
In the [special issue](https://www.degruyter.com/journal/key/jqas/11/1/html) that *JQAS* subsequently published, we learned that the winning entry was based on a fairly simple logistic regression model trained on betting market data [@lopez2015building].
Thus, the competition not only sparked interest in sports analytics, but resulted in peer-reviewed research that, in that case, demonstrated the value of superior data over sophisticated modeling.

Perhaps motivated by his success in the Kaggle March Madness competition, Michael Lopez joined the National Football League and launched the [Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/) in 2019.
This annual competition has similarly fueled new research directions in American football and a [special issue](https://www.degruyter.com/journal/key/jqas/16/2/html) of *JQAS* [@lopez2020bigger].
Successful entries and their corresponding publications [@DeshpandeEvans2020; @Yurko2020going; @ChuReyersThomsonWu2020] have launched the careers of several of the most prominent early-career researchers in sports analytics.

Similar competitions include the [Big Data Cup](https://www.stathletes.com/big-data-cup/) for ice hockey and the [SABR Diamond Dollars Case Competition](https://sabr.org/analytics/case) for baseball.

@@

- How significant is the element of chance in a particular sport? 
- Given that we know who the best teams are, who are the best players and how can we quantify their relative contributions? 
- What combinations of players work best together in a particular sport?
